---
title: "Lab 03"
format: html
editor: visual
---

## Lab 03

Importing Data

```{r}

download.file(
  "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
  destfile = file.path("~", "Downloads", "met_all.gz"),
  method   = "libcurl",
  timeout  = 60
)

library(R.utils)
met <- data.table::fread(file.path("~", "Downloads", "met_all.gz"))
met

```

Finding the Dimensions

```{r}
dim(met)

```

First and Last Rows

```{r}
head(met)
```

```{r}
tail(met)
```

```         

```

Structure of the data

```{r}
str(met)
```

Tabling the data by year

```{r}
table(met$year)
```

Tabling the data by day .

```{r}
table(met$day)
```

Tabling the data by hour.

```{r}
table(met$hour)
```

Summary of Temperature:

```{r}
summary(met$temp)
```

Summary of Elevation:

```{r}
summary(met$elev)
```

Summary of Winds:

```{r}
summary(met$wind.sp)
```

Replacing 9999.0 with NA

```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```

Order Temperature:

```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```

Order Temperature:

```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```

Summary Statistic:

```{r}
elev <- met[elev==max(elev), ]
summary(elev)

```

Correlation Coefficient (wind vs temp):

```{r}
cor(elev$temp, elev$wind.sp, use="complete")
```

Correlation Coefficient (temp vs hour):

```{r}
cor(elev$temp, elev$hour, use="complete")
```

Correlation Coefficient (wind vs day):

```{r}
cor(elev$wind.sp, elev$day, use="complete")

```

Correlation Coefficient (temp vs day)

```{r}
cor(elev$temp, elev$day, use="complete")
```

Histogram of Elevation

```{r}
hist(met$elev, breaks=100)
```

Histogram of Temperatures

```{r}
hist(met$temp)

```

Histogram of Wind Speed

```{r}
hist(met$wind.sp)
```

Mapping

```{r}
#install.packages("leaflet")
library(magrittr)
library(devtools)
library(leaflet)
library(knitr)
#install.packages("knitr")
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>%
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)



```

Look at the time series of temperature and wind speed at this location. For this we will need to create a date-time variable for the x-axis.

```{r}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```

Ordering Date:

```{r}
elev <- elev[order(date)]
head(elev)
```

Plotting Date and Temperature:

```{r}
plot(elev$date, elev$temp, type='l')
```

Plotting Data and Wind:

```{r}
plot(elev$date, elev$wind.sp, type='l')
```

Final Question:

Which day had in August (2018) had the highest temperature?

```{r}
library(ggplot2)

met[met$temp == max(met$temp),]  #The 26th
met[met$temp == min(met$temp),] #The 27th

```

What's the average, max, and median temperature per day?

```{r}
library(dplyr)
tpd<-met %>%group_by(day) %>% summarise(mean(temp), max(temp), median(temp))
tpd

```

Are wind speed and temperature related? Does wind speed depend on temperature?

```{r}
library(dplyr)
cor(elev$temp, elev$wind.sp, use="complete")
met %>% skimr::skim(wind.sp)
met %>% ggplot(aes(x=temp, y=wind.sp)) + geom_jitter() + geom_smooth(method="lm")

lm_1= lm(wind.sp~temp,met )
summary(lm_1)





```
